\LARGE \bf Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model

### 2018-06-09 Conference/ACC 2016/ACC16.1 Old Version/ACC16.1.tex
added Conference folder

\title{\LARGE \bf
Bayesian Occupancy Grid Mapping via Exact Inverse Sensor Model}

@@

Mapping in the context of robotics is to generate a representation of the environment around the robot.
This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment
Various map representations have been applied to the mapping problem such as occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or with distinguishable features in a landmark representation~\cite{MonThrKolWeg02}.
To build maps in real time, a computationally-efficient algorithm is required.% \EditTL{rewrite this paragraph for mapping, without introducing SLAM here}

This paper focusses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells that are considered as either occupied or free. Obtaining the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, known as the \emph{inverse sensor model}, is of fundamental importance with occupancy grid mapping. However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements appear too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.

Instead, several techniques have been proposed to approximate the exact inverse sensor model. %Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
In the early occupancy grid mappings based on sonar sensors, \emph{ad hoc} techniques are employed to obtain an inverse sensor model, where the bearings of depth measurements are considered uncertain because the directions of bouncing sound waves are highly inaccurate~\cite{MorElf85,Elf89}.
%These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
These techniques are based on approximate functions to represent the inverse sensor model with questionable accuracy~\cite{ChoLynHutKanBurKavThr05} and applied to more modern sensors in~\cite{And09},~\cite{PirRutBisSch11}.

The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm, such as neural network to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}. These approaches are undesirable in practice due to complexities associated to implementing a learning algorithm. For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples or how many samples are required to obtain a reasonable approximation. Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.

@@

This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model. More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields a posteriori probability of occupancy of all of the cells within the area covered by the range sensor from the range measurements. The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and it is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model. Furthermore, the proposed approach integrates a priori probability of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps. As such, it is contrast to the existing framework based on log-odds ratios that impose an additional Markov assumption.

@@

Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,...,n_m}$, that is defined as $\mathbf{m}_i=1$ when it is occupied, and $\mathbf{m}_i=0$ when it is free. Another binary random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience.
Therefore, the probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The location and size of each grid cell are assumed to be known.
%, but its occupancy is uncertain; the occupancy is treated as a binary random variable such that $P(\mathbf{m}_i)$ is the probability that $\mathbf{m}_i$ is occupied and $P(\bar{\mathbf{m}}_i)$ is the probability that $\mathbf{m}_i$ is free, such that $P(\mathbf{m}_i)+P(\bar{\mathbf{m}}_i)=1$.
The occupancies of different cells are assumed independent such that
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,...,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
\end{align}

@@

However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem. Early on, the derivation assumes that if $X_t$ and $\mathbf{m}_i$ are known, then $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$, i.e.,
%\begin{align}
%\label{eqn:FirstBayesRule}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})
%&
%=\frac{P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(Z_t|X_{1:t},Z_{1:t-1})}.
%\end{align}
\begin{align}
\label{eqn:AssumptionEarly}
P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t).
\end{align}
The above equation can be justified if both sides are conditioned by the complete map instead of $\mathbf{m}_i$, since the probability distribution of the sensor is completely determined by its pose and the map. However, since it is conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may have certain information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement $Z_t$.  In short, this log-odds framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.

@@

These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.

@@

Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$ are is. Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, a measurement ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.

@@

Next, we compute the occupancy probability of each cell. Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{2_{n_m}}\,|\ \mathbf{m}_i=1\}$. To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e., 
%All possible combinations of map $m$ must be considered in which $\mathbf{m}_i$ is occupied, denoted as $m:m(i)=\mathbf{m}_i$,
\begin{align}
\label{eqn:InvSenModWithProbDens}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align}
Furthermore, to determine the normalizing constant $\eta_{t,l}$, its its complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner. Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, it appears that these equations require additions $2^{n_m}$ times for the $i$-th grid cell, and it should be repeated for other cells.
%However, this is not the case when the bearing of $z_{t,l}$ is known. (\EditTL{Not sure why bearing matters})

\paragraph{Computationally Efficient Approach}
Define a reduced map $r$ where the $k$-th cell is denoted $\mathbf{r}_k$ such that $k\in\braces{1,2,...,n_r}$ with $n_r\leq n_m$.
This reduced map is chosen such that each cell of $r$ corresponds to a grid cell of $m$ that $z_{t,l}$ intersects, ordered by increasing distance, which is easily determined with geometry.
Then let the term $\mathbf{r}_{k+}$ mean that the $k$-th cell is occupied, cells with lower index are free, and cells with greater index may or may not be occupied; put differently, the $k$-th cell $\mathbf{r}_k$ is the closest occupied cell to current pose $X_t$ along reduced map $r$.
This means that the forward sensor model $p(z_{t,l}|\mathbf{r}_{k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell, and \emph{no other cells}.

\begin{prop}
Let the $k$-th unnormalized probability of the inverse sensor model be defined as
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=\frac1{\eta_{t,l}}P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber\\&=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\\
%\label{eqn:UnnormalizedCompl}
%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\frac1{\eta_{t,l}}P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
,
\end{align}
where the normalizer $\eta_{t,l}$ is defined as
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}.
\end{align}
Then, the ray inverse sensor model is computed simply as
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}).
\end{align}
\begin{proof}
See the Appendix 1.
\end{proof}
\end{prop}

%
%%respectively.
%These have the follow solutions that may be obtained in a computationally-efficient manner:
%\begin{align}
%\label{eqn:tildePresult}
%\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
%\\
%\label{eqn:tildePcomplementresult}
%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
%=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
%\end{align}

%
%
%Because $\mathbf{r}_k$ is binary, the following holds:
%\begin{align}
%\label{eqn:allEta}
%\eta_{t,l}&=\frac1{\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})}\nonumber
%\\
%&=
%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
%\end{align}
%where $\eta_{t,l}$ is independent of the index $k$ of the reduced map.
%Rearranging \refeqn{Unnormalized} and \refeqn{UnnormalizedCompl}, the ray inverse sensor model is 
%\begin{align}
%\label{eqn:RayISMAnswer}
%P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\\
%P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\end{align}
%respectively, where $\eta_{t,l}$ comes from \refeqn{allEta} and the unnormalized probabilities come from \refeqn{tildePresult} and \refeqn{tildePcomplementresult}.

Compared with the ray inverse sensor model described with \refeqn{InvSenModWithProbDens} considering $n_r$ possible cells, where the summation should be repeated $2^{n_r}$ times \emph{per cell of the reduced map}, the proposed expressions  \refeqn{allEta} and \refeqn{RayISMAnswer} are \textit{substantially} simpler.
In fact, if the terms inside the summations of \refeqn{Unnormalized} are obtained recursively, only $n_r$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is up to $\mathbf{2^{n_r}}$ \textbf{times faster}.

@@

While the derivation of the complete inverse sensor model is complicated, the algorithm is quite straightforward.
Temporary variables,
\begin{align*}
a_k&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{t,l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{t,l,i},X_t)P(\mathbf{r}_{t,l,i}|X_{1:t-1},Z_{1:t-1}),
\\
b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{t,l,j}|X_{1:t-1},Z_{1:t-1}),
\nonumber\\
c_k&=\sum_{i=k+1}^{n_{r,l}}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_j|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_i,X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}),
\end{align*}
serve to simplify the algorithm such that repeated calculations are removed, and updated occupancy probabilities are obtained recursively.
The pseudocode is summarized in Table \ref{tab:Alg_ISM_2D}.

>>>>

\title{\LARGE \bf
Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model}

@@

Mapping in the context of robotics is to generate a map, which represents the environment in close proximity of a robot.
This problem is crucial to understanding the surrounding regions when a robot enters an uncertain environment.
Various map representations have been applied to the mapping problem such as occupancy grids~\cite{WolSuk05}, octomaps~\cite{WurHorBenStaBur10}, or with distinguishable features in a landmark representation~\cite{MonThrKolWeg02}.
To build maps in real time, a computationally-efficient algorithm is required.% \EditTL{rewrite this paragraph for mapping, without introducing SLAM here}

This paper focusses on the popular occupancy grid representation, where the map space is decomposed into evenly-spaced grid cells that are considered as either occupied or free.
The \emph{inverse sensor model} is the probability of occupancy for a grid cell based on range measurements and the configuration of the mobile robot, which is of fundamental importance with occupancy grid mapping. However, the exact solution to the inverse sensor model has not been utilized in occupancy grid mapping, because the computational requirements appear too cumbersome to compute the exact model in real time~\cite{ThrBurFox05}.

Instead, several techniques have been proposed to approximate the exact inverse sensor model. %Due to the computational limitations, two common approaches are applied to obtain a function close to the inverse sensor model.
In the early occupancy grid mapping applications based on sonar sensors, \emph{ad hoc} techniques are employed to obtain an inverse sensor model, where the bearings of depth measurements are considered uncertain because the directions of bouncing sound waves are highly inaccurate~\cite{MorElf85,Elf89}.
%These occupancy grid mapping schemes employ \emph{ad hoc} techniques to obtain the inverse sensor model.
These techniques are based on approximate functions to represent the inverse sensor model with questionable accuracy~\cite{ChoLynHutKanBurKavThr05} and applied to more modern sensors in~\cite{And09},~\cite{PirRutBisSch11}.

The other popular approach to obtain an approximate inverse sensor model is to simulate maps, robot poses, and measurements and use a learning algorithm, such as a neural network, to approximate the inverse sensor model~\cite{Thr01,ThrBurFox05}. These approaches are undesirable in practice due to complexities associated with implementing a learning algorithm. For example, the accuracy of such inverse sensor models strongly depends on the samples selected for learning, but it is unclear how to select those samples or how many samples are required to obtain a reasonable approximation. Furthermore, it is challenging to apply any learning algorithm over the large dimensional space composed of maps, poses, and measurements.

@@

This paper proposes a computationally efficient algorithm to construct the exact inverse sensor model. More explicitly, for a given forward sensor model defined by the probability distribution of range measurements, this algorithm yields a posterior probability of occupancy of all of the cells within the area covered by the range sensor from the range measurements. The key idea is reducing the computational load by using the fact that if a cell is occupied, the occupancy of the other cells blocked by it is irrelevant to the forward sensor model, and this model is systematically utilized with various probabilistic properties to derive a computationally-efficient solution to the inverse sensor model. Furthermore, the proposed approach integrates a priori probability of occupancy and multiple range measurements according to the Bayesian framework to obtain more accurate maps. As such, the proposed approach contrasts from the existing framework based on log-odds ratios that impose an additional Markov assumption.

@@

Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,...,n_m}$, that is defined as $\mathbf{m}_i=1$ when it is occupied, and $\mathbf{m}_i=0$ when it is free. Another binary random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience.
Therefore, the probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The location and size of each grid cell is assumed to be known.
%, but its occupancy is uncertain; the occupancy is treated as a binary random variable such that $P(\mathbf{m}_i)$ is the probability that $\mathbf{m}_i$ is occupied and $P(\bar{\mathbf{m}}_i)$ is the probability that $\mathbf{m}_i$ is free, such that $P(\mathbf{m}_i)+P(\bar{\mathbf{m}}_i)=1$.
The occupancies of different cells are assumed independent such that
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,...,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
\end{align}

@@

However, the log-odds ratio formulation makes an assumption that is not consistent with the occupancy grid mapping problem. Early on, the derivation assumes that if $X_t$ and $\mathbf{m}_i$ are known, then $Z_t$ is independent of $X_{1:t-1}$ and $Z_{1:t-1}$, i.e.,
%\begin{align}
%\label{eqn:FirstBayesRule}
%P(\mathbf{m}_i|X_{1:t},Z_{1:t})
%&
%=\frac{P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})P(\mathbf{m}_i|X_{1:t},Z_{1:t-1})}{P(Z_t|X_{1:t},Z_{1:t-1})}.
%\end{align}
\begin{align}
\label{eqn:AssumptionEarly}
P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t).
\end{align}
The above equation can be justified if both sides are conditioned by the complete map instead of $\mathbf{m}_i$, since the probability distribution of the sensor is completely determined by its pose and the map. However, since it is conditioned only by the $i$-th cell, past poses $X_{1:t-1}$ and past measurement scans $Z_{1:t-1}$ may have certain information about the occupancy of the other cells, which should be considered in calculating the probability distribution of the measurement scan $Z_t$.  In short, this log-odds framework is based on the assumption that neglects potentially important information contained in $X_{1:t-1}$ and $Z_{1:t-1}$.

@@

%These are contrary to inverse sensor models commonly used with the log-odds ratio formulations, and the proposed algorithm integrates all of past poses and measurement scans with the current ones to obtain more accurate estimate of the map.

@@

Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$, is known. Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, a measurement ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.

@@

Next, we compute the occupancy probability of each cell. Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{2_{n_m}}\,|\ \mathbf{m}_i=1\}$. To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e., 
%All possible combinations of map $m$ must be considered in which $\mathbf{m}_i$ is occupied, denoted as $m:m(i)=\mathbf{m}_i$,
\begin{align}
\label{eqn:InvSenModWithProbDens}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align}
Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner. Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, it appears that these equations require additions $2^{n_m}$ times for the $i$-th grid cell, and it should be repeated for other cells.
%However, this is not the case when the bearing of $z_{t,l}$ is known. (\EditTL{Not sure why bearing matters})

\paragraph{Computationally Efficient Approach}
Define a reduced map $r$ where the $k$-th cell is denoted $\mathbf{r}_k$ such that $k\in\braces{1,2,...,n_r}$ with $n_r\leq n_m$.
This reduced map is chosen such that each cell of $r$ corresponds to a grid cell of $m$ that $z_{t,l}$ intersects, ordered by increasing distance, which is easily determined with geometry.
Then let the term $\mathbf{r}_{l,k+}$ mean that the $k$-th cell of the $l$-th ray is occupied, cells with lower index are free, and cells with greater index may or may not be occupied; put differently, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along reduced map $r$.
This means that the forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell, and \emph{no other cells}.

\begin{prop}
For the $l$-th measurement ray, let the $k$-th unnormalized probability of the inverse sensor model be defined as
\begin{align}
\label{eqn:Unnormalized}
\tilde P&(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%=\frac1{\eta_{t,l}}P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})
\nonumber\\&=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
%\\
%\label{eqn:UnnormalizedCompl}
%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\frac1{\eta_{t,l}}P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})
.
\end{align}
Then, the ray inverse sensor model is computed simply as
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
\end{align}
where the normalizer $\eta_{t,l}$ is defined as
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1}.
\end{align}
\begin{proof}
See the Appendix 1.
\end{proof}
\end{prop}

%
%%respectively.
%These have the follow solutions that may be obtained in a computationally-efficient manner:
%\begin{align}
%\label{eqn:tildePresult}
%\tilde P(\mathbf{r}_k|z_{t,l},&X_{1:t},Z_{1:t-1})=P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\left\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{k+},X_t)P(\mathbf{r}_k|X_{1:t-1},Z_{1:t-1}),
%\\
%\label{eqn:tildePcomplementresult}
%\tilde P(\bar{\mathbf{r}}_k|z_{t,l},&X_{1:t},Z_{1:t-1})
%=P(\bar{\mathbf{r}}_k|X_{1:t-1},Z_{1:t-1})\nonumber\\
%&\quad\times \sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})
%\nonumber
%\\
%&\quad
%+
%\sum_{i=k+1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\
%&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1}).
%\end{align}

%
%
%Because $\mathbf{r}_k$ is binary, the following holds:
%\begin{align}
%\label{eqn:allEta}
%\eta_{t,l}&=\frac1{\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})+\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})}\nonumber
%\\
%&=
%\bigg[\sum_{i=1}^{n_r}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{i+},X_t)P(\mathbf{r}_i|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
%\end{align}
%where $\eta_{t,l}$ is independent of the index $k$ of the reduced map.
%Rearranging \refeqn{Unnormalized} and \refeqn{UnnormalizedCompl}, the ray inverse sensor model is 
%\begin{align}
%\label{eqn:RayISMAnswer}
%P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\\
%P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\bar{\mathbf{r}}_k|z_{t,l},X_{1:t},Z_{1:t-1}),
%\end{align}
%respectively, where $\eta_{t,l}$ comes from \refeqn{allEta} and the unnormalized probabilities come from \refeqn{tildePresult} and \refeqn{tildePcomplementresult}.

Compared with the ray inverse sensor model described with \refeqn{InvSenModWithProbDens} considering $n_r$ possible cells, where the summation should be repeated $2^{n_r}$ times \emph{per cell of the reduced map}, the proposed expressions  \refeqn{allEta} and \refeqn{RayISMAnswer} are \textit{substantially} simpler.
In fact, if the terms inside the summations of \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, only $n_r$ repetitions are required for \emph{all cells of the reduced map}, yielding an algorithm that is up to $\mathbf{2^{n_r}}$ \textbf{times faster}.

@@

While the derivation of the complete inverse sensor model is complicated, the algorithm is quite straightforward.
Temporary variables,
\begin{align*}
a_k&=\sum_{i=1}^{k-1}\left\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
\\
b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1}),
\nonumber\\
c_k&=\sum_{i=k+1}^{n_{r,l}}\left\{\prod_{j=0}^{i-2}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\right\}\nonumber\\&\quad\quad\times p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
\end{align*}
serve to simplify the algorithm such that repeated calculations are removed, and updated occupancy probabilities are obtained recursively.
The pseudocode is summarized in Table \ref{tab:Alg_ISM_2D}.
